\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing income inequality with tax data - Switzerland from 1917 to 2010}

### 2015-01-09 paper/subdocuments/results.Rnw
1993 rein, 1995 raus.

<<different_ginis, fig.cap='Gini coefficients for different income measures',fig.pos='H'>>=
df$G_taxed[df$G_taxed==1]<-NA
library(reshape)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
names(dflong)[3]<-"Income"
names(dflong)[4]<-"Gini"
levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=bemessungsjahre,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.28,0.55))+
  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Income))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Income definitions within tax data")
fig_incomes
# was ist die maximale Bandbreite zwischen den Punkten?
max(filter(df,kanton=="CH") %>% mutate(d_steink_reink=abs(G_steink-G_reink),d_steink_taxed=abs(G_steink-G_taxed),d_reink_taxed=abs(G_reink-G_taxed)) %>% select(starts_with("d_")),na.rm=TRUE)
# 0.026

@@

<<setup_data_Brühlhart with and without taxed>>=
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
bd<-subset(bd,(Jahr<=1995 | Jahr >2002)&Veranlagungsperiode!=1995)
# 2011 zufügen
bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)%>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
levels(bd$Veranlagungsperiode)<- c(levels(bd$Veranlagungsperiode),"2011")
bd <- rbind(bd,bd.2011[,names(bd)])
bd.long<-reshape(bd,
varying = c("gini_steink","gini_steinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("taxable income (only taxed)","equivalised taxable income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
bd.0<-subset(bd.0,(Jahr==1995|Jahr>=2003)&Veranlagungsperiode!=1995)
# 2011 zufügen
bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011) %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
bd.0 <- rbind(bd.0,bd.0.2011[,names(bd.0)])
bd.0.long<-reshape(bd.0,
varying = c("gini_steink","gini_steinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("taxable income","equivalised taxable income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[3]<-"Income"
@

@@

<<with_without_zeros_Gini,warning=FALSE, fig.cap='Gini coefficients with, without and imputed non-taxed units',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
#library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="including non-taxed"]<-NA
dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="imputed values for non-taxed"]<-NA
fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
       aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.28,0.55))+
  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Influence of non-taxed")
fig_zeros

@@

estv_masse <- estv %>% 
  filter(kanton=="CH",
         eink_art=="Nach steuerbarem Einkommen",
         fall=="Normalfälle",
         anz_pflichtige>0) %>%
        # steuerperiode!=1945.5) %>% 
  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
  group_by(factor(bemessungsjahre)) %>%
  mutate(mean_steink=steink/anz_pflichtige) %>%
  summarise(Gini=gini(mean_steink,weights=anz_pflichtige),Atkinson=getAtkinson(mean_steink,w=anz_pflichtige),Theil=getTheil(mean_steink,w=anz_pflichtige))
estv_masse$id<-1
names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")
estv_masse<-estv_masse[as.numeric(estv_masse$Year)>4,]
try(detach("package:reshape", unload=TRUE))
library(reshape2)
estv_masse_long <- melt(estv_masse,id=c("id","Year"))

estv_masse_long$measure <- factor(estv_masse_long$variable)
estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
estv_masse_long <- estv_masse_long %>% group_by(measure) %>% mutate(log_value=log(value)-log(value)[1])
fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
  scale_y_continuous(limits=c(-0.7,0.45))+
  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=measure),size=3)+
  geom_line(aes(linetype=measure))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using different inequality measures")
fig_measures

>>>>

<<different_ginis, fig.cap='Gini coefficients for different income measures',fig.pos='H'>>=
df$G_taxed[df$G_taxed==1]<-NA
library(reshape)
df <- df %>% filter(kanton=="CH"&(bemessungsjahre<=1993.5|bemessungsjahre>=2003))
number_ticks <- function(n) {function(limits) pretty(limits, n)}
dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
names(dflong)[3]<-"Income"
names(dflong)[4]<-"Gini"
levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=bemessungsjahre,y=Gini,shape=Income))+
  scale_y_continuous(limits=c(0.28,0.55))+
  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=Income),size=3)+
  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Income))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Income definitions within tax data")
fig_incomes
# was ist die maximale Bandbreite zwischen den Punkten?
max(filter(df,kanton=="CH") %>% mutate(d_steink_reink=abs(G_steink-G_reink),d_steink_taxed=abs(G_steink-G_taxed),d_reink_taxed=abs(G_reink-G_taxed)) %>% select(starts_with("d_")),na.rm=TRUE)
# 0.026

@@

<<setup_data_Brühlhart with and without taxed>>=
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
bd<-subset(bd,(Jahr<=1993 | Jahr >2002)&Veranlagungsperiode!=1995)
# 2011 zufügen
bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)%>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
levels(bd$Veranlagungsperiode)<- c(levels(bd$Veranlagungsperiode),"2011")
bd <- rbind(bd,bd.2011[,names(bd)])
bd.long<-reshape(bd,
varying = c("gini_steink","gini_steinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("taxable income (only taxed)","equivalised taxable income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %>% 
  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
bd.0<-subset(bd.0,(Jahr<=1993|Jahr>=2003)&Veranlagungsperiode!=1995)
# 2011 zufügen
bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011) %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
bd.0 <- rbind(bd.0,bd.0.2011[,names(bd.0)])
bd.0.long<-reshape(bd.0,
varying = c("gini_steink","gini_steinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("taxable income","equivalised taxable income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
names(bdlong)[3]<-"Income"
@

@@

<<with_without_zeros_Gini,warning=FALSE, fig.cap='Gini coefficients with, without and imputed non-taxed units',fig.pos='H'>>=
df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
#library(reshape)
dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
#dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre!=1993.5&dflong$bemessungsjahre<2003&dflong$Type=="including non-taxed"]<-NA
#dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre!=1993.5&dflong$bemessungsjahre<2003&dflong$Type=="imputed values for non-taxed"]<-NA
fig_zeros <- ggplot(filter(dflong,kanton=="CH"&
  (((bemessungsjahre<=1993.5|bemessungsjahre>=2003)&Type=="without non-taxed")|
  ((bemessungsjahre==1993.5|bemessungsjahre>=2003)&Type!="without non-taxed"))), 
       aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
  scale_y_continuous(limits=c(0.28,0.55))+
  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
  geom_line(size=0.5)+
  geom_point(aes(shape=Type),size=3)+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Influence of non-taxed")
fig_zeros

@@

estv_masse <- estv %>% 
  filter(kanton=="CH",
         eink_art=="Nach steuerbarem Einkommen",
         bemessungsjahre<=1993.5|bemessungsjahre>=2003,
         fall=="Normalfälle",
         anz_pflichtige>0) %>%
        # steuerperiode!=1945.5) %>% 
  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
  group_by(factor(bemessungsjahre)) %>%
  mutate(mean_steink=steink/anz_pflichtige) %>%
  summarise(Gini=gini(mean_steink,weights=anz_pflichtige),Atkinson=getAtkinson(mean_steink,w=anz_pflichtige),Theil=getTheil(mean_steink,w=anz_pflichtige))
estv_masse$id<-1
names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")
estv_masse<-estv_masse[as.numeric(estv_masse$Year)>4,]
try(detach("package:reshape", unload=TRUE))
library(reshape2)
estv_masse_long <- melt(estv_masse,id=c("id","Year"))

estv_masse_long$Measure <- factor(estv_masse_long$variable)
estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
estv_masse_long <- estv_masse_long %>% group_by(Measure) %>% mutate(log_value=log(value)-log(value)[1])
fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=Measure))+
  scale_y_continuous(limits=c(-0.7,0.45))+
  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
  geom_point(aes(shape=Measure),size=3)+
  geom_line(aes(linetype=Measure))+
  theme_bw()+
  xlab("Year")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
  ggtitle("Using different inequality measures")
fig_measures
